Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-25 Manuscript/manuscript.Rmd
edits

Given that many TFBS are known to colocalize together on the linear genome and are therefore highly correlated, we sought to avoid the incidence of overfitting our optimally built classifier by performing recursive feature elimination (Supplementary Figure 9) to establish the necessary number of TFBS to include. We found that our cell line specific model was able to obtain near optimal performance using only ~8 TFBS (Figure 4B; Supplementary Figure 8B).

<!--
## PREDICTIVE IMPORTANCES CONFIRMED BIOLOGICAL FUNCTIONALITY OF KNOWN TRANSCRIPTION FACTOR BINDING SITES

We then sought to determine which of the transcription factor binding sites were most influential in predicting TAD boundaries. We found that the cell line specific binding sites associated with SMC3, CTCF, RAD21, and ZNF143 were given consistently high predictive importance, across all resolutions (Figure 6 C). The SMC3, CTCF, and RAD21 binding sites are known factors associated with the cohesin protein complex, which is hypothesized to be what creates loops in the DNA  [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likewise, ZNF143 is a known structural regulators of enhancer-promoter interactions that also contribute to looping of the linear genome [@weintraub2017yy1; @gonzalez2017znf143]. With the identification of these factors known to be enriched at TAD boundaries, we are able to confirm the validity of our model building process. Additionally, we were also able to identify some lesser known cell line specific transcription factors as predictive of TAD boundaries including the MYC-associated zing-finger protein (MAZ), jun-D proto-oncogene (JUND), and ELF1. These transcription factor binding sites tended to be given higher predictive importance at lower resolutions. [??? Where do you show it?] In fact, at lower resolutions, the predictive importance was found to be much more evenly dispersed across the top transcription factor binding sites, than for higher resolutions. This is again suggestive of the fact that the spatial relationship between functional annotations, specifically transcription factor binding sites, and TAD boundaries changes at different Hi-C data resolutions.

![**Assessing optimally built cell line specific random forest classifiers for predicting TAD boundary formation.** (A) Comparing the proportion of total predictive importance carried by each group of functional genomic annotation across resolution. Predictive importance was averaged for models built on chromosome specific regions of the linear genome. Proportional importance was then calculated according to the annotation group of the feature space including chromatin state segmentations (CSS), histone modifications (HM), and transcription factor binding sites (TFBS). (B) Comparing models using specific groups of functional genomic annotation as the feature space in addition to a model with a feature space containing all 3 types combined (combined; purple). Error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosomal section of the linear genome. (C) Heatmaps illustrating the importance of the top 20 cell line specific transcription factors across resolution. Importances were sorted according to the row-wise average across resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red and dark blue for the GM12878 and K562 cell line respectively.](Figures/prop_imps_a_comparing_annots_b_var_imps_c_v2.png)

-->

>>>>

Given that many TFBS are known to colocalize together on the linear genome and are therefore highly correlated, we sought to avoid the incidence of overfitting our optimally built classifier by performing recursive feature elimination (Supplementary Figure 9) to establish the necessary number of TFBS to include. We found that our cell line specific model was able to obtain near optimal performance using only ~8 TFBS (Figure 4B; Supplementary Figure 8B). However, given that we are modeling on chromosome specific sections of the genome, these 8 annotations are not the same for all chromosomes. To determine the most common TFBS, we clustered the predictive importances across chromosomes (Figure 4C; Supplementary Figure 8C). We found that the following transcription factors were most important in predicting TAD boundary regions: SMC3, ZNF143, CTCF, and RAD21. These are known architectural proteins responsible for the 3D structure of the human genome. Namely, the SMC3, CTCF, and RAD21 binding sites are known factors associated with the cohesin protein complex, which is hypothesized to be what creates loops in the DNA  [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likewise, ZNF143 is a known structural regulators of enhancer-promoter interactions that also contribute to looping of the linear genome [@weintraub2017yy1; @gonzalez2017znf143].

![**Assessing top ranked transcription factors in TAD boundary prediction.** (A) (B) (C)](Figures/annot_specific_rfe_varimps_heatmap.png)
